### 2014-08-02 Pero_Transcriptome.tex
working on results

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that only identical sequences were removed. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

Transcriptome assembly was accomplished using the program Trinity. The raw assembly contained 743314 assembled sequenced measuring 418Mb. This assembly was filtered using TMP \textgreater 1 as a threshold. The filtered assembly contained 130764 sequences measuring 149Mb. from this filteres dataset, I extracted 64355 putative coding sequences (60Mb). Of these 64355 sequences, 37960 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (16880 sequences), 3-prime end (4203 sequences), or were internal (5312 sequencing having neither stop nor start codon).

>>>>

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

Transcriptome assembly for each tissue type was accomplished using the program Trinity. The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively. This assembly was filtered using a blast procedure with resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts. Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome. Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.

I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues.

After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained. From this filtered concatenated dataset, I extracted 64355 putative coding sequences (60Mb). Of these 64355 sequences, 37960 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (16880 sequences), 3-prime end (4203 sequences), or were internal (5312 sequencing having neither stop nor start codon).
